A weighted data representation based approach for manifold learning

### 2018-11-04 DimReduction/Lle/Documentation/Article_lle.tex
Finishing theorical framework in Lle documentation

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
%\usepackage{babelbib}

@@

\begin{abstract}
    We present a detailed implementation of the local linear embeding
    algorithm from both a theorical and practical approximation. In this work the
    dimensionality reduction is achieved by a two step procedure. First a
    weighted data representation based in $K$ nearest neighbors is calculated, and
    then a minimization of distance between the data representation and a low
    dimensionality configuration is conducted. Finally the resulting algorithm
    is tested on 3-dimensional manifolds.
\end{abstract}

@@

An approximation to solve this problems are dimensionality reduction techniques.
These methods provide a way to discover compact representations of
multidimensional datasets in a low dimensional space. This approaches allow the
analysis process to be carried out in an easier way and help to reduce the
probability of getting erroneous results.

@@

However, there are still some issues related to the usage of these methods.
Assumptions like the use of euclidean distances as a measure in the
multidimensional space implies that the data is homogeneous or isotropic, and
this is really often not the case in many real world datasets
\cite{homogeneous1}, \cite{homogeneous2}. Because of this, the usage of other
distance metrics is a good option to increase the performance of these approches
\cite{dist1}, \cite{dist2}.

@@

An appropriate processing method for dimentionality reduction in the case of
a nonlinear approach can be manifold learning techniques \cite{manifold1},
\cite{manifold2}. This algorithms seek to generate nonlinear dimensionality
reductions by exploiting

% missing part
\cite{lle}

In this work the linear local embeding algorithm is used to perform a
dimensionality reduction of a multidimentional dataset by a manifold learning
technique. The first part of the work will carafully describe each part of the
mathemathical backgound of the method, then the implementation details of the
algorithm are shown to prepare a results section, and finally some conclusions
are exposed.

@@

    As mention before, the problem that we are trying to solve is to find
    a low dimensional representation of points in a manifold that is in a high
    dimensional space. In order to achieve this, two steps are considered. The
    first one is to find a way to represent each point in the original dataset as
    a linear combination of its $K$ nearest neighbors. This means that the
    objective is to find a set of weights that minimize the following
    expression

@@

    This problem can be solved by optimizing an arbitrary case $i$. The first
    step is to define the neighborhood matrix of $x_i$, a vector that will
    conatin the weights of the combination and a special matrix $e$ that will
    facilitate the process later.

@@

    At this point, if $\lambda$ is wrongly chosen, then the result would be a
    scaled version of the true value. Then, $\lambda$ can be chosen
    arbitrarily. Then, after $W_i$ is computed a rescale procedure can be
    conducted in order to satisfy the restriction.\\

    A good choice for the variable is $\lambda = 2$. This leads to the
    cancellation of the fraction and reduces the problem to find the solution of
    a linear system of equations

@@

    This procedure is used to find all the weight values. Then this weights are
    used to define a configuration of points in a low dimensional space that
    is as similar as possible to the original dataset. The objective in this
    point of the problem is minimize the following expression

@@

    To solve this problem define

    \[
        W =
        \begin{bmatrix}
            |   &  |  &       & |   \\
            W_1 & W_2 & \dots & W_n \\
            |   &  |  &       & |
        \end{bmatrix}_{n \times n}
        where
        \hspace{0.5cm}
        W_i =
        \begin{bmatrix}
            0 \\
            w_1 \\
            \vdots \\
            w_k \\
            0 \\
        \end{bmatrix}_{n \times 1}
    \]

    Where $W_i$ is a vector that contains the value of the weight that correspond
    to the $j$ neighbor of $x_i$ in the corresponding position, and 0 anywhere
    else.\\

@@

    \begin{equation*}
        \sum_{j=1}^k w_{ij} N(y_i)_j = Y W_i
    \end{equation*}

    Which means that the problem can be rewrited as

@@

    Now,

@@

    Which is a vector that has a $1$ in the $i$ position and $0$ anywhere else.
    With this new vector we can redefine the problem as

@@

    So, this optimization problem can be solved as follows

@@

    \begin{equation*}
        \begin{aligned}
            \frac{\delta \mathcal{L}}{\delta W_i} &= Y M^t + Y M - 2 \lambda Y\\
                                                  &= Y M + Y M - 2 \lambda Y\\
                                                  &= 2 Y M - 2 \lambda Y \\
                              2 Y M - 2 \lambda Y &= 0 \\
                                            2 Y M &= 2 \lambda Y \\
                                              Y M &= \lambda Y \\
                                              (Y M)^t &= (\lambda Y)^t \\
                                              M^t Y^t &= Y^t \lambda \\
                                              M Y^t &= \lambda Y^t \\
        \end{aligned}
    \end{equation*}

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}

@@

\begin{abstract}
    We present a detailed implementation of the local linear embeding algorithm
    from both a theoretical and practical approximation. In this work the
    dimensionality reduction is achieved by a two step procedure. First a
    weighted data representation based in $K$ nearest neighbors is calculated,
    and then a minimization of distance between the data representation and a
    low dimensionality configuration is conducted. Finally the resulting
    algorithm is tested on 3-dimensional manifolds.
\end{abstract}

@@

An approximation to solve these problems are dimensionality reduction techniques.
These methods provide a way to discover compact representations of
multidimensional datasets in a low dimensional space. This approaches allow the
analysis process to be carried out in an easier way and help to reduce the
probability of getting erroneous results.

@@

However, there are still some issues related to the usage of these methods.
Assumptions like the use of euclidean distances as a measure in the
multidimensional space, implies that the data is homogeneous or isotropic, and
this is really often not the case in many real world datasets
\cite{homogeneous1}, \cite{homogeneous2}. Because of this, the usage of other
distance metrics is a good option to increase the performance of these approches
\cite{dist1}, \cite{dist2}.

@@

An appropriate processing method for dimentionality reduction in the case of
a nonlinear approach can be manifold learning techniques \cite{manifold1},
\cite{manifold2}. This algorithms seek to generate nonlinear dimensionality
reductions by assuming that the data points are samples from a low dimensional
structure which is embedded in a multidimensional space and hence a simpler
representation can be constructed. \cite{manifold}

One example of a manifold learning algorithm is Local Linear Embeding (LLE)
\cite{lle}. This technique uses the fact that from a local perspective a
non-linear structure can be aproximated with an euclidean measure. Hence an
aproximation of a multidimensional manifold in a low dimensional space can be
conducted by representing small pieces, where the euclidean assumption can hold,
and then using these pieces together to get a reconstruction as similar as 
possible to the original manifold.

The objective of this work is to build a theoretical and practical develop of
the ideas used in the LLE algorithm to build a manifold learning method. For
this purpose the first part of the work will carafully describe each part of
the mathemathical backgound of the method, then the implementation details of
the algorithm are shown to prepare a results section, and finally some
conclusions are exposed.

@@

    As mention before, the problem that we are trying to solve is to find
    a low dimensional representation of points in a manifold that is in a high
    dimensional space. A theorical approximation of 
    the method is developed taking as staring point the description given in
    \cite{proof}. In this method two steps are considered. The
    first one is to find a way to represent each point in the original dataset as
    a linear combination of its $K$ nearest neighbors. This means that the
    objective is to find a set of weights that minimize the following
    expression

@@

    This problem can be solved by optimizing an arbitrary case $i$. The first
    step is to define the neighborhood matrix of $x_i$, also a vector that will
    conatin the weights of the combination, and finally a special matrix $e$ that
    is filled with ones.

@@

    At this point, if $\lambda$ is wrongly chosen, then the result would be a
    scaled version of the optimal value. Then, $\lambda$ can be chosen
    arbitrarily and after $W_i$ is computed, a rescale procedure can be
    conducted in order to satisfy the restriction.\\

    Hence, a good choice for the variable is $\lambda = 2$. This leads to the
    cancellation of the fraction and reduces the problem to find the solution of
    a linear system of equations

@@

    This procedure is used to find all of the weights values. Then these weights
    are used to define a configuration of points in a low dimensional space that
    is as similar as possible to the original dataset. The objective in this
    point is to minimize the following expression

@@

    To solve this problem, define

    \[
        \hspace{-0.5cm}
        W =
        \begin{bmatrix}
            |   &  |  &       & |   \\
            W_1 & W_2 & \dots & W_n \\
            |   &  |  &       & |
        \end{bmatrix}_{n \times n}
        where
        \hspace{0.25cm}
        W_i =
        \begin{bmatrix}
            0 \\
            w_1 \\
            \vdots \\
            w_k \\
            0 \\
        \end{bmatrix}_{n \times 1}
    \]

    Here $W_i$ is a vector that contains the value of the weight that correspond
    to the $j$ neighbor of $x_i$ in the corresponding position, and 0 anywhere
    else.\\

@@

    \begin{equation*}
        \sum_{j=1}^k w_{ij} N(y_i)_j = Y W_i
    \end{equation*}\\

    Which means that the problem can be rewritten as

@@

    Now, just as before, the objective is to express $y_i$ as vectors operations.
    To achieve this, define

@@

    Which is a vector that has a $1$ in the $i$ position and $0$ anywhere else.
    With this new vector the problem can redefined as

@@

    This optimization problem can be solved as follows

@@

    \begin{equation*}
        \begin{aligned}
            \frac{\delta \mathcal{L}}{\delta W_i} &= Y M^t + Y M - 2 \lambda Y\\
                                                  &= Y M + Y M - 2 \lambda Y\\
                                                  &= 2 Y M - 2 \lambda Y
        \end{aligned}
    \end{equation*}

    Finally, by setting this derivative to 0 the solution of the problem can
    be founded

    \begin{equation*}
        \begin{aligned}
            2 Y M - 2 \lambda Y &= 0 \\
                          2 Y M &= 2 \lambda Y \\
                            Y M &= \lambda Y \\
                            (Y M)^t &= (\lambda Y)^t \\
                            M^t Y^t &= Y^t \lambda \\
                            M Y^t &= \lambda Y^t \\
        \end{aligned}
    \end{equation*}

    Which is a eigenvector problem and can be solved with classical methods.

@@

\cite{matlab}
